---
title: "House Prices"
# TODO: Añadid vuestro nombres
author: "Javier Herrer Torres"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    toc: true
    toc_depth: 3
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
house_prices <- read.csv("train.csv")
par(mfrow = c(1, 1))
```

## Data preparation

First, the training data was imported through the `read.csv` function.

Then, 10 factors are selected using the continuos description method and filtering by the 10 most related factors to the target.
Before that, factors should have the appropiate type.
The selected factors are:

1. overall material and finish of the house,
1. physical locations within the Ames city limits,
1. quality of the material on the exterior,
1. basement height evaluation,
1. kitchen quality,
1. interior finish of the garage,
1. fireplace quality,
1. type of foundation,
1. garage location,
1. type of dwelling involved in the sale.

```{r}
library(FactoMineR)

house_prices$MSSubClass <- factor(house_prices$MSSubClass)
house_prices$MSZoning <- factor(house_prices$MSZoning)
house_prices$Street <- factor(house_prices$Street)
house_prices$Alley <- factor(house_prices$Alley)
house_prices$LotShape <- factor(house_prices$LotShape)
house_prices$LandContour <- factor(house_prices$LandContour)
house_prices$Utilities <- factor(house_prices$Utilities)
house_prices$LotConfig <- factor(house_prices$LotConfig)
house_prices$LandSlope <- factor(house_prices$LandSlop)
house_prices$Neighborhood <- factor(house_prices$Neighborhood)
house_prices$Condition1  <- factor(house_prices$Condition1)
house_prices$Condition2 <- factor(house_prices$Condition2)
house_prices$BldgType <- factor(house_prices$BldgType)
house_prices$HouseStyle <- factor(house_prices$HouseStyle)
house_prices$OverallQual <- factor(house_prices$OverallQual)
house_prices$OverallCond <- factor(house_prices$OverallCond)
house_prices$RoofStyle <- factor(house_prices$RoofStyle)
house_prices$RoofMatl <- factor(house_prices$RoofMatl)
house_prices$Exterior1st <- factor(house_prices$Exterior1st)
house_prices$Exterior2nd <- factor(house_prices$Exterior2nd)
house_prices$MasVnrType <- factor(house_prices$MasVnrType)
house_prices$ExterQual <- factor(house_prices$ExterQual)
house_prices$ExterCond <- factor(house_prices$ExterCond)
house_prices$Foundation <- factor(house_prices$Foundation)
house_prices$BsmtCond <- factor(house_prices$BsmtCond)
house_prices$BsmtExposure <- factor(house_prices$BsmtExposure)
house_prices$BsmtFinType1 <- factor(house_prices$BsmtFinType1)
house_prices$BsmtFinType2 <- factor(house_prices$BsmtFinType2)
house_prices$Heating <- factor(house_prices$Heating)
house_prices$HeatingQC <- factor(house_prices$HeatingQC)
house_prices$CentralAir <- factor(house_prices$CentralAir)
house_prices$Electrical <- factor(house_prices$Electrical)
house_prices$KitchenQual <- factor(house_prices$KitchenQual)
house_prices$Functional <- factor(house_prices$Functional)
house_prices$FireplaceQu <- factor(house_prices$FireplaceQu)
house_prices$GarageFinish <- factor(house_prices$GarageFinish)
house_prices$GarageQual <- factor(house_prices$GarageQual)
house_prices$Heating <- factor(house_prices$Heating)
house_prices$GarageCond <- factor(house_prices$GarageCond)
house_prices$PavedDrive <- factor(house_prices$PavedDrive)
house_prices$PoolQC <- factor(house_prices$PoolQC)
house_prices$Fence <- factor(house_prices$Fence)
house_prices$MiscFeature <- factor(house_prices$MiscFeature)
house_prices$SaleType <- factor(house_prices$SaleType)
house_prices$SaleCondition <- factor(house_prices$SaleCondition)

continuos_description <- condes(house_prices, 81)
continuos_description$quali

relevant_factors <- rownames(continuos_description$quali[1:10, ])
numeric_variables <- sapply(house_prices, is.numeric)

house_prices <- cbind(
  house_prices[, numeric_variables],
  house_prices[, relevant_factors]
)
```








